Supporting Information:\\Black-swan events in animal populations

### 2016-11-24 text/supp.tex
Add random walk model to supplement and add missing bst file
Simplify supplement for submission

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{verbose, letterpaper, tmargin = 2.54cm, bmargin = 2.54cm,
  lmargin = 2.54cm, rmargin = 2.54cm}
\geometry{letterpaper}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{ragged2e}
\usepackage{verbatim}
\usepackage[sectionbib]{chapterbib}
\usepackage{amsmath}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{array}
\usepackage{tabu} % for spacing between rows in longtable
\usepackage{titlesec}
\usepackage{url}
\usepackage{mathtools}
%\usepackage{cite}
% \usepackage{citesupernumber}
% \usepackage{scicite}

\usepackage{lineno}
\usepackage{xcolor}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\modulolinenumbers[2]

% Linux Libertine:
\usepackage{textcomp}
\usepackage[sb]{libertine}
\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
\usepackage[cal=boondoxo]{mathalfa} % mathcal
%\useosf % osf for text, not math
\usepackage[supstfm=libertinesups,%
  supscaled=1.2,%
  raised=-.13em]{superiors}

@@

% Fix line numbering and align environment
% http://phaseportrait.blogspot.ca/2007/08/lineno-and-amsmath-compatibility.html
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
  \renewenvironment{#1}%
     {\linenomath\csname old#1\endcsname}%
     {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
  \patchAmsMathEnvironmentForLineno{#1}%
  \patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

@@

\author{
Sean C. Anderson$^{1*}$ \and
Trevor A. Branch$^2$ \and
Andrew B. Cooper$^3$ \and
Nicholas K. Dulvy$^1$
}
\date{}

@@

\textsuperscript{*}Corresponding author: Sean C. Anderson; E-mail: sandrsn@uw.edu; Present address: School of Aquatic and Fishery Sciences, Box 355020, University of Washington, Seattle, WA 98195, USA

\linenumbers
\onehalfspacing

@@

\subsection{Density independence} We fit a simplified version of the
Gompertz model in which the density dependence parameter \(b\) was fixed at
\(1\) (density independent). This is equivalent to fitting a random walk model
(with drift) to the \(\log\) abundances or assuming the growth rates are drawn
from a stationary distribution. The model was as follows:
\begin{align}
x_t &= \lambda + x_{t-1} + \epsilon_t\\
\epsilon &\sim \mathrm{Student}\mhyphen t(\nu, 0, \sigma).
\end{align}
We fit this model for three reasons: (1) it is computationally simpler and so
provides a check that our more complicated full Gompertz model was obtaining
reasonable estimates of \(\nu\), (2) it provides a test of whether density
dependence was affecting our perception of heavy tails, (3) it
matches how some previous authors have modelled heavy tails without accounting
for density dependence \citep{segura2013}.

@@

\section{Code availability.} Data and code to reproduce our analysis are
available at \url{https://github.com/seananderson/heavy-tails} (currently
private).

@@

\bibliographystyle{ecologyletters}\bibliography{jshort,newbib}

@@

%\begin{helvetica}
\singlespacing
\begin{scriptsize}
\begin{longtable}{>{\RaggedRight}m{1.4cm}>{\RaggedRight}p{6.0cm}>{\RaggedRight}p{0.7cm}>{\RaggedRight}p{1.2cm}>{\RaggedRight}p{4.4cm}>{\RaggedRight}p{1.4cm}}
\caption{Populations with a high probability of heavy-tailed dynamics in the base heavy-tailed Gompertz population dynamics model. Shown are the log abundance time series, population descriptions, Global Population Dynamics Database Main IDs, citation for the data source or separate verification literature, a description of the cause of the black-swan events (if known), the probability of heavy tails as calculated by our model, and median estimate of $\nu$ from our model with 90\% quantile credible intervals indicated in parentheses. Red dots on the time series indicate downward black-swan events and blue values indicate upward black-swan events that have a $10^{-4}$ probability or less of occurring if the population dynamics were explained by a Gompertz model with normally distributed process noise with a standard deviation equal to the scale parameter in the fitted $t$ distribution.}\\

@@

@@

\noindent
Example Stan code for a heavy-tailed Gompertz model. The specific code for
the various models in our analysis is available at
\url{https://github.com/seananderson/heavy-tails} (currently private).

>>>>

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{verbose, letterpaper, tmargin = 2.54cm, bmargin = 2.54cm,
  lmargin = 2.54cm, rmargin = 2.54cm}
\geometry{letterpaper}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{ragged2e}
\usepackage{verbatim}
% \usepackage[sectionbib]{chapterbib}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{array}
% \usepackage{tabu} % for spacing between rows in longtable
% \usepackage{titlesec}
\usepackage{url}
\usepackage{mathtools}
%\usepackage{cite}
% \usepackage{citesupernumber}
% \usepackage{scicite}

% \usepackage{lineno}
% \usepackage{xcolor}
% \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
% \modulolinenumbers[2]

% Linux Libertine:
% \usepackage{textcomp}
% \usepackage[sb]{libertine}
% \usepackage[varqu,varl]{inconsolata}% sans serif typewriter
% \usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
% \usepackage[cal=boondoxo]{mathalfa} % mathcal
% %\useosf % osf for text, not math
% \usepackage[supstfm=libertinesups,%
%   supscaled=1.2,%
%   raised=-.13em]{superiors}

@@

% % Fix line numbering and align environment
% % http://phaseportrait.blogspot.ca/2007/08/lineno-and-amsmath-compatibility.html
% \newcommand*\patchAmsMathEnvironmentForLineno[1]{%
%   \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
%   \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
%   \renewenvironment{#1}%
%      {\linenomath\csname old#1\endcsname}%
%      {\csname oldend#1\endcsname\endlinenomath}}%
% \newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
%   \patchAmsMathEnvironmentForLineno{#1}%
%   \patchAmsMathEnvironmentForLineno{#1*}}%
% \AtBeginDocument{%
% \patchBothAmsMathEnvironmentsForLineno{equation}%
% \patchBothAmsMathEnvironmentsForLineno{align}%
% \patchBothAmsMathEnvironmentsForLineno{flalign}%
% \patchBothAmsMathEnvironmentsForLineno{alignat}%
% \patchBothAmsMathEnvironmentsForLineno{gather}%
% \patchBothAmsMathEnvironmentsForLineno{multline}%
% }

@@

\author{
Sean C. Anderson$^{1,2*}$ \and
Trevor A. Branch$^2$ \and
Andrew B. Cooper$^3$ \and
Nicholas K. Dulvy$^1$
}
\date{}

@@

\textsuperscript{*}Corresponding author: Sean C. Anderson; E-mail: sandrsn@uw.edu

% \linenumbers
\onehalfspacing

@@

\subsection{Density independence} We fit a simplified version of the
Gompertz model in which the density dependence parameter \(b\) was fixed at
\(1\) (density independent). This is equivalent to fitting a random walk model
(with drift) to the \(\log\) abundances or assuming the growth rates are drawn
from a stationary distribution. The model was as follows:
\begin{align}
  \label{eq:rw-drift}
x_t &= \lambda + x_{t-1} + \epsilon_t\\
\epsilon &\sim \mathrm{Student}\mhyphen t(\nu, 0, \sigma).
\end{align}
We fit this model for three reasons: (1) it is computationally simpler and so
provides a check that our more complicated full Gompertz model was obtaining
reasonable estimates of \(\nu\), (2) it provides a test of whether density
dependence was affecting our perception of heavy tails, (3) it
matches how some previous authors have modelled heavy tails without accounting
for density dependence \citep{segura2013}.

We also considered a simpler model: a random walk on log abundance without drift,
\begin{align}
x_t &= x_{t-1} + \epsilon_t\\
\epsilon &\sim \mathrm{Student}\mhyphen t(\nu, 0, \sigma).
\end{align}
However, the estimates of $\nu$ were qualitatively similar to those from
Equation \ref{eq:rw-drift} (Fig.~\ref{fig:alt}C) and are therefore not shown.

@@

\section{Code availability.} Data and code to reproduce our analysis are
available at \url{https://github.com/seananderson/heavy-tails}.

@@

\bibliographystyle{apalike}\bibliography{jshort,newbib}

@@

%\begin{helvetica}
\singlespacing
\begin{scriptsize}
\begin{longtable}{>{\RaggedRight}m{1.4cm}>{\RaggedRight}p{5.0cm}>{\RaggedRight}p{0.65cm}>{\RaggedRight}p{2.1cm}>{\RaggedRight}p{4.2cm}>{\RaggedRight}p{1.6cm}}
\caption{Populations with a high probability of heavy-tailed dynamics in the base heavy-tailed Gompertz population dynamics model. Shown are the log abundance time series, population descriptions, Global Population Dynamics Database Main IDs, citation for the data source or separate verification literature, a description of the cause of the black-swan events (if known), the probability of heavy tails as calculated by our model, and median estimate of $\nu$ from our model with 90\% quantile credible intervals indicated in parentheses. Red dots on the time series indicate downward black-swan events and blue values indicate upward black-swan events that have a $10^{-4}$ probability or less of occurring if the population dynamics were explained by a Gompertz model with normally distributed process noise with a standard deviation equal to the scale parameter in the fitted $t$ distribution.}\\

@@

% \clearpage

@@

\noindent
Example Stan code for a heavy-tailed Gompertz model. The specific code for
the various models in our analysis is available at
\url{https://github.com/seananderson/heavy-tails}.
